---
title: "Madison Common Council Voting Record"
format: 
  html:
    toc: true
    toc-location: left
---

```{r}
#| label: loaddata
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(kableExtra)
library(purrr)

load("alder_data_processed.RData")

```

## % YIMBY Voting Record of Current Alders, 2023-25 term

```{r}
#| label: tbl-YIMBY_percentage
#| tbl-cap: Percentage of contested votes by active alders that were pro-housing (YIMBY).
#| echo: false
#| warning: false
#| message: false
kableExtra::kable(YIMBY.prop, "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))



```

## All Housing Votes, 2023-Present


### 2025

```{r}
#| label: tbl-all-votes-2025
#| tbl-cap: Contested votes of active alders in 2025, listed by their district. YIMBY votes in blue, anti-YIMBY votes in red.
#| echo: false
#| warning: false
#| message: false

#Only keep votes in YIMBY.vote.list from 2025
YIMBY.vote.list.2025 <- YIMBY.vote.list %>%
  filter(year(date) == 2025) %>%
  select(-date)

    kable(YIMBY.vote.list.2025, "html", escape = FALSE) %>%
  column_spec(6,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "N"))
                               | is.na(YIMBY.vote.list$`1`)  , "blue", "red")) %>%
  column_spec(7,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "N"))
                               | is.na(YIMBY.vote.list$`2`)  , "blue", "red")) %>%
  column_spec(8,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "N"))
                               | is.na(YIMBY.vote.list$`3`)  , "blue", "red")) %>%
 column_spec(9,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "N"))
                               | is.na(YIMBY.vote.list$`4`)  , "blue", "red")) %>%
  column_spec(10,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "Y") |
                             (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "N"))
                           | is.na(YIMBY.vote.list$`5`)  , "blue", "red")) %>%
  column_spec(11,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "N"))
                               | is.na(YIMBY.vote.list$`6`)  , "blue", "red")) %>%
  column_spec(12,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "N"))
                               | is.na(YIMBY.vote.list$`7`)  , "blue", "red")) %>%
  column_spec(13,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "N"))
                               | is.na(YIMBY.vote.list$`8`)  , "blue", "red")) %>%
  column_spec(14,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "N"))
                               | is.na(YIMBY.vote.list$`9`)  , "blue", "red")) %>%
        column_spec(15,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "N"))
                               | is.na(YIMBY.vote.list$`10`)  , "blue", "red")) %>%
        column_spec(16,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "N"))
                               | is.na(YIMBY.vote.list$`11`)  , "blue", "red")) %>%
        column_spec(17,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "N"))
                               | is.na(YIMBY.vote.list$`12`)  , "blue", "red")) %>%
        column_spec(18,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "N"))
                               | is.na(YIMBY.vote.list$`13`)  , "blue", "red")) %>%
        column_spec(19,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "N"))
                               | is.na(YIMBY.vote.list$`14`)  , "blue", "red")) %>%
        column_spec(20,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "N"))
                               | is.na(YIMBY.vote.list$`15`)  , "blue", "red")) %>%
        column_spec(21,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "N"))
                               | is.na(YIMBY.vote.list$`16`)  , "blue", "red")) %>%
        column_spec(22,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "N"))
                               | is.na(YIMBY.vote.list$`17`)  , "blue", "red")) %>%
        column_spec(23,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "N"))
                               | is.na(YIMBY.vote.list$`18`)  , "blue", "red")) %>%
       column_spec(24,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "N"))
                               | is.na(YIMBY.vote.list$`19`)  , "blue", "red")) %>%
       column_spec(25,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "N"))
                               | is.na(YIMBY.vote.list$`20`)  , "blue", "red")) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))



```





### 2024

```{r}
#| label: tbl-all-votes-2024
#| tbl-cap: Contested votes of active alders in 2024, listed by their district. YIMBY votes in blue, anti-YIMBY votes in red.
#| echo: false
#| warning: false
#| message: false

#Only keep votes in YIMBY.vote.list from 2024
YIMBY.vote.list.2024 <- YIMBY.vote.list %>%
  filter(year(date) == 2024) %>%
  select(-date)

    kable(YIMBY.vote.list.2024, "html", escape = FALSE) %>%
  column_spec(6,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "N"))
                               | is.na(YIMBY.vote.list$`1`)  , "blue", "red")) %>%
  column_spec(7,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "N"))
                               | is.na(YIMBY.vote.list$`2`)  , "blue", "red")) %>%
  column_spec(8,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "N"))
                               | is.na(YIMBY.vote.list$`3`)  , "blue", "red")) %>%
 column_spec(9,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "N"))
                               | is.na(YIMBY.vote.list$`4`)  , "blue", "red")) %>%
  column_spec(10,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "Y") |
                             (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "N"))
                           | is.na(YIMBY.vote.list$`5`)  , "blue", "red")) %>%
  column_spec(11,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "N"))
                               | is.na(YIMBY.vote.list$`6`)  , "blue", "red")) %>%
  column_spec(12,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "N"))
                               | is.na(YIMBY.vote.list$`7`)  , "blue", "red")) %>%
  column_spec(13,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "N"))
                               | is.na(YIMBY.vote.list$`8`)  , "blue", "red")) %>%
  column_spec(14,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "N"))
                               | is.na(YIMBY.vote.list$`9`)  , "blue", "red")) %>%
        column_spec(15,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "N"))
                               | is.na(YIMBY.vote.list$`10`)  , "blue", "red")) %>%
        column_spec(16,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "N"))
                               | is.na(YIMBY.vote.list$`11`)  , "blue", "red")) %>%
        column_spec(17,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "N"))
                               | is.na(YIMBY.vote.list$`12`)  , "blue", "red")) %>%
        column_spec(18,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "N"))
                               | is.na(YIMBY.vote.list$`13`)  , "blue", "red")) %>%
        column_spec(19,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "N"))
                               | is.na(YIMBY.vote.list$`14`)  , "blue", "red")) %>%
        column_spec(20,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "N"))
                               | is.na(YIMBY.vote.list$`15`)  , "blue", "red")) %>%
        column_spec(21,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "N"))
                               | is.na(YIMBY.vote.list$`16`)  , "blue", "red")) %>%
        column_spec(22,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "N"))
                               | is.na(YIMBY.vote.list$`17`)  , "blue", "red")) %>%
        column_spec(23,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "N"))
                               | is.na(YIMBY.vote.list$`18`)  , "blue", "red")) %>%
       column_spec(24,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "N"))
                               | is.na(YIMBY.vote.list$`19`)  , "blue", "red")) %>%
       column_spec(25,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "N"))
                               | is.na(YIMBY.vote.list$`20`)  , "blue", "red")) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))



```





### 2023

```{r}
#| label: tbl-all-votes-2023
#| tbl-cap: Contested votes of active alders in 2023, listed by their district. YIMBY votes in blue, anti-YIMBY votes in red.
#| echo: false
#| warning: false
#| message: false

#Only keep votes in YIMBY.vote.list from 2023
YIMBY.vote.list.2023 <- YIMBY.vote.list %>%
  filter(year(date) == 2023) %>%
  select(-date)

    kable(YIMBY.vote.list.2023, "html", escape = FALSE) %>%
  column_spec(6,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`1` == "N"))
                               | is.na(YIMBY.vote.list$`1`)  , "blue", "red")) %>%
  column_spec(7,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`2` == "N"))
                               | is.na(YIMBY.vote.list$`2`)  , "blue", "red")) %>%
  column_spec(8,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`3` == "N"))
                               | is.na(YIMBY.vote.list$`3`)  , "blue", "red")) %>%
 column_spec(9,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`4` == "N"))
                               | is.na(YIMBY.vote.list$`4`)  , "blue", "red")) %>%
  column_spec(10,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "Y") |
                             (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`5` == "N"))
                           | is.na(YIMBY.vote.list$`5`)  , "blue", "red")) %>%
  column_spec(11,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`6` == "N"))
                               | is.na(YIMBY.vote.list$`6`)  , "blue", "red")) %>%
  column_spec(12,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`7` == "N"))
                               | is.na(YIMBY.vote.list$`7`)  , "blue", "red")) %>%
  column_spec(13,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`8` == "N"))
                               | is.na(YIMBY.vote.list$`8`)  , "blue", "red")) %>%
  column_spec(14,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`9` == "N"))
                               | is.na(YIMBY.vote.list$`9`)  , "blue", "red")) %>%
        column_spec(15,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`10` == "N"))
                               | is.na(YIMBY.vote.list$`10`)  , "blue", "red")) %>%
        column_spec(16,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`11` == "N"))
                               | is.na(YIMBY.vote.list$`11`)  , "blue", "red")) %>%
        column_spec(17,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`12` == "N"))
                               | is.na(YIMBY.vote.list$`12`)  , "blue", "red")) %>%
        column_spec(18,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`13` == "N"))
                               | is.na(YIMBY.vote.list$`13`)  , "blue", "red")) %>%
        column_spec(19,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`14` == "N"))
                               | is.na(YIMBY.vote.list$`14`)  , "blue", "red")) %>%
        column_spec(20,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`15` == "N"))
                               | is.na(YIMBY.vote.list$`15`)  , "blue", "red")) %>%
        column_spec(21,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`16` == "N"))
                               | is.na(YIMBY.vote.list$`16`)  , "blue", "red")) %>%
        column_spec(22,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`17` == "N"))
                               | is.na(YIMBY.vote.list$`17`)  , "blue", "red")) %>%
        column_spec(23,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`18` == "N"))
                               | is.na(YIMBY.vote.list$`18`)  , "blue", "red")) %>%
       column_spec(24,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`19` == "N"))
                               | is.na(YIMBY.vote.list$`19`)  , "blue", "red")) %>%
       column_spec(25,color = ifelse(((YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "Y") |
                                 (!YIMBY.vote.list$YIMBY & YIMBY.vote.list$`20` == "N"))
                               | is.na(YIMBY.vote.list$`20`)  , "blue", "red")) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))



```






## Methodology

The "YIMBY" ("Yes in my backyard") voting scale is intended to be an objective measure of the pro-housing alignment of Madison Common Council Members. However, you should be highly skeptical of any measure that purports to be "objective," because every scale of attitudes requires some value judgement. Our measure is no different! When evaluating a vote we employ maximalist YIMBY criteria, even though this is not necessarily how we would vote and is not the view of any particular pro-housing advocates or organization. But it provides a useful reference point. The standard holds that housing developments should always be approved if they meet minimal standards for safety and functionality; zoning that prevents the most housing possible from being built should be eliminated; transportation policy should promote car-free living (or at least the ability to live without a car). All alder votes relevant to YIMBY issues are weighted equally, with votes on particular amendments or motions of an item averaged together so that an item doesn't get more weight because it had multiple votes for amendments, etc. 

Do you think this is a misguided approach? Please send us an email and let us know why we are wrong. Also please let us know if you think we are missing votes or have made a mistake.


